### 2017-07-14 methyl-valerate.tex
Updates

\documentclass[letterpaper, review, sort&compress]{elsarticle}

\journal{Fuel}

\usepackage{multirow}

\usepackage{lineno}

\usepackage{enumitem}
\setlist{noitemsep}

\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\sisetup{
    separate-uncertainty=true,
    per-mode=symbol
}

\usepackage{graphicx}
\usepackage{pgf}
\usepgflibrary{fpu}

\usepackage[tableposition=top]{caption}
\usepackage{booktabs}
\usepackage{mathtools}

\newcommand*{\logten}{\mathop{\log_{10}}}

\DeclareSIUnit\torr{Torr}
\DeclareSIUnit\atm{atm}

\bibliographystyle{elsarticle-num-CNF}
\usepackage{doi}
\usepackage[capitalize]{cleveref}
\begin{document}
%======================================================================
\begin{frontmatter}
\title{ Experiments and Modeling of Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }

\author[uconn]{Bryan W.\ Weber\corref{cor1}}
\author[uconn]{Justin A.\ Bunnell}
\author[idaho]{Kamal Kumar}
\author[uconn]{Chih-Jen Sung}
\address[uconn]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
\address[idaho]{Department of Mechanical Engineering, University of Idaho, Moscow, ID, USA}
\cortext[cor1]{Corresponding Author: bryan.weber@uconn.edu}

%====================================================================
\begin{abstract} % not to exceed 200 words

    Methyl valerate (\ce{C6H12O2}, methyl pentanoate) is a methyl ester and a relevant surrogate
    component for biodiesel. In this work, we present ignition delays of methyl valerate measured
    using a rapid compression machine at a range of engine-relevant temperature, pressure, and
    equivalence ratio conditions. The conditions we have studied include equivalence ratios
    \((\phi)\) from \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and pressures
    of \SIlist{15;30}{\bar}. The ignition delay data demonstrate a negative temperature coefficient
    region in the temperature range of \SIrange[range-phrase={--}]{720}{800}{\K} for both
    \(\phi=2.0\), \SI{15}{\bar} and \(\phi=1.0\), \SI{30}{\bar}, with two-stage ignition apparent
    over the narrower temperature ranges of \SIrange[range-phrase={--}]{720}{760}{\K} for
    \SI{15}{\bar} and \SIrange[range-phrase={--}]{740}{760}{\K} at \SI{30}{\bar}. In addition, the
    experimental ignition delay data are compared with simulations using an existing chemical
    kinetic model from the literature. The simulations with the literature model under-predict the
    data by factors between \numlist{2;10} over the entire range of the experimental data. In
    addition, a new chemical kinetic model is developed using the Reaction Mechanism Generator (RMG)
    software. The agreement between the experimental data and the RMG model is also not
    satisfactory. To help determine the possible reasons for the disagreement, a path analysis of
    both models is completed. It is found that improvements to both the reaction pathways and
    thermodynamic properties are required. Further directions for future improvement of the methyl
    valerate model are discussed.

\end{abstract}

% (Provide 2-4 keywords describing your research. Only abbreviations firmly
% established in the field may be used. These keywords will be used for
% sessioning/indexing purposes.)
\begin{keyword}
    chemical kinetics\sep rapid compression machine\sep autoignition\sep methyl ester\sep methyl valerate\sep methyl pentanoate
\end{keyword}
\end{frontmatter}
%======================================================================

\linenumbers

\section{Introduction}\label{sec:introduction}

For transportation applications, biodiesel is an important constituent in improving environmental
friendliness of fuels. This is due to its renewability when produced from sustainable agricultural
crops and its ability to reduce emissions relative to petroleum-derived fuels~\cite{Hoekman2012}.
Biodiesel typically consists of long-chain methyl ester molecules, with typical compositions of
\ce{C14} to \ce{C20}~\cite{Hoekman2012}. Recognizing that the large molecular size of the methyl
esters within biodiesel fuel makes creating and using detailed chemical kinetic models
challenging~\cite{Lai2011}, it is desired to study their combustion chemistry by studying simpler
methyl ester molecules.

A recent review paper summarizes the work on methyl esters relevant to biodiesel
combustion~\cite{Coniglio2013}; the following summary focuses on ignition delay measurements, since
these are the focus of this paper. Autoignition of methyl butanoate (MB, \ce{C5H10O2}) has been
well-studied in both shock tube and rapid compression machine experiments~
\cite{Metcalfe2007,Walton2009,Dooley2008,Akih-Kumgeh2010,Akih-Kumgeh2011,Hadj-Ali2009,Kumar2016}.
The prevalence of MB data in the literature is largely due to the early identification of MB as a
potential surrogate fuel for biodiesel~\cite{Fisher2000}. However, the experiments have shown that
MB may not be an appropriate surrogate for biodiesel, due to its lack of negative temperature
coefficient (NTC) behavior, a requirement for a suitable biodiesel surrogate~\cite{Coniglio2013}.

Methyl esters larger than MB, such as methyl valerate (MV, \ce{C6H12O2}, methyl pentanoate), have
also been studied as possible biodiesel surrogates. \citet{Hadj-Ali2009} used a rapid compression
machine (RCM) to study the autoignition of several methyl esters including MV. Although MV exhibited
two-stage ignition in this study, little additional research has been done on its low-temperature
chemistry. \citet{Korobeinichev2015} studied MV in premixed laminar flames and extended a detailed
high temperature chemical kinetic model to include MV and methyl hexanoate. \citet{Dmitriev2015}
added MV to n-heptane/toluene fuel blends to determine the resulting intermediate species in
premixed flames using a flat burner at \SI{1}{atm} and an equivalence ratio of 1.75. The addition of
MV helped reduce soot forming intermediates including benzene, cyclopentadienyl, acetylene,
propargyl, and vinylacetylene~\cite{Dmitriev2015}. \citet{Hayes2009} computationally examined the
peroxy radical isomerization reactions for MV to better understand the low temperature reaction
pathways. Finally, \citet{Dievart2013} used diffusion flames in the counterflow configuration to
determine extinction limits for a number of methyl esters, including MV, and validated a detailed
kinetic model with the experimental data.

This work provides additional data for the autoignition of MV. Data is collected in a RCM under
engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios \((\phi)\) from
\numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
out to provide additional information on the fidelity of using MV as a biodiesel surrogate.

\section{Experimental Methods}\label{sec:experimental-methods}

The RCM used in this study is a single piston arrangement and is pneumatically driven and
hydraulically stopped. The device has been described in detail previously~\cite{Mittal2007a} and
will be described here briefly for reference. The end of compression (EOC) temperature and pressure
(\(T_C\) and \(P_C\) respectively), are independently changed by varying the overall compression
ratio, initial pressure \((P_0)\), initial temperature \((T_0)\), and specific heat ratio of the
experiments. The piston in the reaction chamber is machined with a specially designed crevice to
suppress the roll-up vortex effect and promote homogeneous conditions in the reactor during and
after compression~\cite{Mittal2006}.

The primary diagnostic on the RCM is the in-cylinder pressure measured by a Kistler 6125C dynamic
transducer that is compensated for thermal shock. The transducer is coupled to a Kistler 5010B
charge amplifier. The voltage output of the charge amplifier is recorded by a National Instruments
9125 analog input device connected to a cDAQ 9178 chassis. The voltage is sampled at a rate of
either 50 kHz or 100 kHz by a LabView VI and processed by a Python package called
UConnRCMPy~\cite{Weber2016a}. Version 3.0.5 of UConnRCMPy~\cite{uconnrcmpy}, 3.6 of Python, 2.3.0
of Cantera~\cite{cantera}, 1.13 of NumPy~\cite{vanderWalt2011}, 0.19.0 of SciPy~\cite{Jones2001},
and 2.0.1 of Matplotlib~\cite{Hunter2007} are used in the analysis in this paper.

\begin{center}
    \captionsetup{type=figure}
    \includegraphics{figures/ignition_delay_definition.pdf}
    \caption{Definition of the ignition delays used in this work. The
    experiment in this figure is conducted for a \(\phi=2.0\) mixture with
    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{Ar})=0.5\)}, \(P_0=\SI{0.7806}{\bar}\),
    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
    \(\tau=\SI{27.56\pm0.89}{\ms}\), and \(\tau_1=\SI{16.60\pm0.46}{\ms}\).}
    \label{fig:ign-delay-def}
\end{center}

The compression stroke of the RCM brings the fuel/oxidizer mixture to the EOC conditions, and for
suitable thermodynamic states, the mixture will ignite after a delay period. The definitions of the
ignition delays are shown in \cref{fig:ign-delay-def}. The time of the EOC is defined as the maximum
of the pressure trace prior to the start of ignition and the ignition delays are defined as the time
from the EOC until local maxima in the first time derivative of the pressure. Each experimental
condition is repeated at least five times to ensure repeatability of the data. As there is some
random scatter present in the data, the standard deviation \((\sigma)\) of the ignition delays from
the runs at a given condition is computed. Typically, \(\sigma\) is less than \SI{10}{\percent} of
the mean values of the overall ignition delay \((\tau)\) and the first stage ignition delay
\((\tau_1)\).

In addition to the reactive experiments, non-reactive experiments are conducted by replacing \ce{O2}
with \ce{N2} to determine the influence of machine-specific behavior on the experimental conditions
and permit the calculation of the EOC temperature via the isentropic relations between pressure and
temperature~\cite{Lee1998}. The EOC temperature is calculated by the procedure described in
\cref{sec:computational-methods}.

The mixtures considered in this study are shown in \cref{tab:mixtures}. Four equivalence ratios of
MV in ``air'' are considered. While \ce{O2} is kept at \SI{21}{\percent} by mole in the oxidizer,
the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer is varied to adjust the temperatures reached at the
EOC. Two \(P_C\) conditions are studied in this work, \SIlist{15;30}{\bar}, representing
engine-relevant conditions. For the \(\phi=2.0\) condition, only \(P_C = \SI{15}{\bar}\) is
considered because we could not achieve \(T_C\) values low enough that the ignition during the
compression stroke can be prevented.

\begin{center}
    \captionsetup{type=table}
    \caption{Mixtures considered in this work}
    \begin{tabular}{Sccccc}
        \toprule
        {\(\phi\)} & \multicolumn{4}{c}{Mole Fraction (purity)} & \ce{Ar}/(\ce{N2} + \ce{Ar}) \\
        \cmidrule{2-5}
         & \ce{MV} (\SI{100}{\percent}) & \ce{O2} (\SI{99.994}{\percent}) & \ce{Ar} (\SI{99.999}{\percent}) & \ce{N2} (\SI{99.999}{\percent}) &  \\
        \midrule
        0.25 & 0.0065 & 0.2087 & 0.7848 & 0.0000 & 1.0 \\
        0.5 & 0.0130 & 0.2074 & 0.7796 & 0.0000 & 1.0 \\
        1.0 & 0.0256 & 0.2047 & 0.7697 & 0.0000 & 1.0 \\
        1.0 & 0.0256 & 0.2047 & 0.3849 & 0.3848 & 0.5 \\
        2.0 & 0.0499 & 0.1996 & 0.0000 & 0.7505 & 0.0 \\
        2.0 & 0.0499 & 0.1996 & 0.3752 & 0.3753 & 0.5 \\
        \bottomrule
    \end{tabular}
    \label{tab:mixtures}
\end{center}

Mixtures are prepared in stainless steel mixing tanks, \SI{17}{\L} and \SI{15}{\L} in size. The
proportions of reactants in the mixture are determined by specifying the absolute mass of the fuel,
the equivalence ratio, and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are
made by first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV
is a liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
gravimetrically to within \SI{0.01}{\g} of the specified value. The fuel is injected into the mixing
tank through a septum. Proportions of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room
temperature and the total pressure is measured by an Omega Engineering MMA100V10T2D0T4A6 type static
pressure transducer. The same transducer is used to measure the pressure of the reactants prior to
an experiment.

The RCM is equipped with heaters to control the initial temperature of the mixture. After filling in
the components to the mixing tanks, the heaters are switched on and the system is allowed
\SI{1.5}{\hour} to come to steady state. The mixing tanks are also equipped with magnetic stir bars
so the reactants are well mixed for the duration of the experiments.

\begin{center}
    \captionsetup{type=figure}
    \input{figures/vapor_pressure.pgf}
    \caption{Saturated vapor pressure of MV as a function of temperature,
    plotted using the Antoine equation, \cref{eq:antoine}, with
    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
    \label{fig:vapor-pressure}
\end{center}

The initial temperature is chosen such that the saturated vapor pressure \((P_{\text{sat}})\) of the
fuel at the initial temperature is at least twice the partial pressure of the fuel in the mixing
tank. The Antoine equation
%
\begin{equation}\label{eq:antoine}
    \logten{P_{\text{sat}}} = A - \frac{B}{T - C}
\end{equation}
%
is used to model the saturated vapor pressure of MV as a function of temperature \((T)\), where
\(A\), \(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are
given in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and~\citet{Stephenson1987}.
Unfortunately, the values of the coefficients are different among all three references. Therefore,
coefficients for use in \cref{eq:antoine} are determined in this work by least squares fitting of
the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and \citet{Verevkin2008} using the
\verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure}
shows that the coefficients fitted with this procedure give good agreement with the experimental
data; values for the coefficients computed in this work and in the literature works are given in
\cref{tab:antoine}. The data used to calculate the coefficients are provided in the Supplementary
Material.

\begin{center}
    \captionsetup{type=table}
    \caption{Antoine Equation coefficients computed in this work and from the
    literature. The $2\sigma$ confidence interval is estimated by taking the
    square root of the diagonals of the covariance matrix returned from
    \texttt{curve\_fit()}}
    \label{tab:antoine}
    \begin{tabular}{cccccc}
        \toprule
        & $A$ & $B$ & $C$ & $T_{\text{min}}$, \si{\K} & $T_{\text{max}}$, \si{\K} \\
        \midrule
        This Work & 6.4030 & 1528.69 & 52.881 & 274.9 & 417.18 \\
        $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
        \citet{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
        \citet{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
        \citet{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
        \bottomrule
    \end{tabular}
\end{center}

\section{Computational Methods}\label{sec:computational-methods}
\subsection{RCM Modeling}\label{sec:experimental-modeling}

The Python 3.6 interface of Cantera~\cite{cantera} version 2.3.0 is used for all simulations in this
work. Detailed descriptions of the use of Cantera for these simulations can be found in the work
of~\citet{Weber2016a} and~ \citet{Dames2016}; a brief overview is given here. As mentioned in
\cref{sec:experimental-methods}, non-reactive experiments are conducted to characterize the
machine-specific effects on the experimental conditions in the RCM. This pressure trace is combined
with the reactive pressure trace and used to compute a volume trace by assuming that the reactants
undergo a reversible, adiabatic, constant composition (i.e., isentropic) compression during the
compression stroke and an isentropic expansion after the EOC. The volume trace is applied to a
simulation conducted in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delays from the simulations are
defined in the same manner as in the experiments. The time derivative of the pressure in the
simulation is computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}.

To the best of our knowledge, there are three mechanisms for MV combustion available in the
literature. The first two, by \citet{Korobeinichev2015} and \citet{Dmitriev2015}, were developed to
simulate flames, and do not include the low-temperature chemistry necessary to simulate the
conditions in these experiments. The third model was developed by \citet{Dievart2013} and includes
low-temperature chemistry of MV, although it was only validated by comparison with flame extinction
limits. In converting this mechanism for use in Cantera, we found that there were many species in
the thermodynamic database with multiple data entries. For most of these species the thermodynamic
data is identical. However, some species are not exact duplicates. For these species, it is not
clear from the thermodynamic database file which data set should be preferred. Since Cantera (and
CHEMKIN) choose the first instance of a duplicate species to be used, we retained the first entry
for all duplicated species. The detailed model of \citet{Dievart2013} includes 1105 species and 7141
reactions, and the Cantera formatted input file is available in the Supplementary Material.

\subsection{Reaction Mechanism Generator}\label{sec:reaction-mechanism-generator}

In addition to using a mechanism from the literature, we investigate the use of an automatic
mechanism generator, the open-source Reaction Mechanism Generator (RMG)~\cite{Allen2012} version
2.1.0. The Python version of RMG is used, which requires Python 2.7, and version 2.1.0 of the RMG
database is used. The final RMG model contains 427 species and 13640 reactions. Note that the number
of species is much lower than the \citet{Dievart2013} model because the RMG model focuses on only
one fuel (MV), but the number of reactions is substantially higher. The input file used to generate
the model is available in the Supplementary Material. In addition, the CHEMKIN and Cantera formatted
input files for the RMG model are available in the Supplementary Material.

\section{Experimental Results}\label{sec:experimental-results}
\subsection{Ignition Delays}\label{sec:ignition-delays}

\Cref{fig:ignition-delays} shows the ignition delay results measured in this study. Filled markers
denote the overall ignition delay and hollow markers indicate the first-stage ignition delay.
Vertical error bars are drawn on the symbols to represent the \(2\sigma\) uncertainty in the
ignition delay; for many of the experiments, the uncertainty is approximately the same size as the
data point, so the error bar is hidden. Horizontal error bars are shown on the first and last points
of each equivalence ratio indicating the estimated uncertainty in the EOC temperature of
$\pm\SI{1}{\percent}$~\cite{Weber2015}. \cref{fig:ignition-delays}a shows the results for a
compressed pressure of \SI{15}{\bar}, while \cref{fig:ignition-delays}b shows the results for a
compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not collected for
\SI{30}{\bar}, so there are no red triangle data points in \cref{fig:ignition-delays}b. All the data
are available in comma-separated value files and ChemKED-format \cite{Weber2017} files in the
Supplementary Material.

It can be seen from \cref{fig:ignition-delays} that the ignition delays for the \(\phi=0.25\text{
and } 0.5\) mixtures do not show an NTC region of the ignition delay for both of the pressures
studied in this work. However, the \(\phi=1.0\) mixture shows an NTC region at
\(P_C=\SI[number-unit-product={\ }]{30}{\bar}\) between approximately \SIlist{720;800}{\K}, with
measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the \(\phi=2.0\) mixture
shows an NTC region of ignition delay at \SI{15}{\bar} from approximately \SIrange{720}{780}{\K},
with measured first-stage ignition delays between \SIlist{720;750}{\K}.

\begin{center}
    \captionsetup{type=figure}
    \includegraphics{figures/ignition-delays.pdf}
    \caption{Ignition delays of MV as a function of inverse temperature for varying equivalence
    ratios. Filled points are the overall ignition delays and hollow points are the first stage
    ignition delays. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
    \label{fig:ignition-delays}
\end{center}

\citet{Hadj-Ali2009} also observed two-stage ignition of MV in stoichiometric mixtures, stating that
``[m]ethyl pentanoate... was more reactive [than methyl butanoate] with a limit below which
autoignition no longer occurs observed at \(T_c = \SI{670}{\K}\) and \(P_c = \SI{11.4}{\bar}\). At
this temperature, the autoignition occurred in two stages with a clearly identified cool flame
event.'' However, we do not find two stage ignition for the similar pressure of
\(P_C = \SI{15}{\bar}\) in this study. We note that the stated temperature of the experiment from
the work of \citet{Hadj-Ali2009} (\SI{670}{\K}) is much lower than the lowest temperature we
considered in this work at \SI{15}{\bar}, \(\phi=1.0\) (\SI{828}{\K}). We did not conduct
experiments at lower temperatures because the work of \citet{Mittal2006} showed that the temperature
field in the RCM reaction chamber was uniform for approximately \SI{100}{\ms} after the EOC, and our
measured ignition delay at \SI{15}{\bar}, \(\phi=1.0\), and \SI{828}{\K} is \SI{92.14}{\ms}.

However, we note NTC behavior of the ignition delay and two-stage ignition at the higher pressure of
\SI{30}{\bar}, and at higher temperatures than those reported for two-stage ignition in the study of
\citet{Hadj-Ali2009}. The trend of NTC behavior shifting to higher temperatures with increasing
pressure can be seen in other classes of fuels. \citet{Kukkadapu2012} found a similar trend in
gasoline composed of iso-alkanes, n-alkanes, cyclo-alkanes, aromatics, and olefins.
\citet{Kukkadapu2012} attributed the shift of the NTC region to the reactions between the
hydroperoxyalkyl radical (\ce{QOOH}) and \ce{O2} becoming more dominant than the unimolecular
decomposition of \ce{QOOH} at higher pressures. Similar trends could occur for the hydroperoxyalkyl
radicals of MV.

To further understand the effect of the methyl ester functional group on the NTC region of ignition
delay, we compare with the alkane and alcohol with 5-carbon alkyl chains, n-pentane and n-pentanol.
n-Pentane and MV have the same fuel mole percentage for stoichiometric mixtures in air
(\SI{2.56}{\percent}), while n-pentanol has a fuel mole percentage of \SI{2.72}{\percent} for
stoichiometric conditions. \citet{Ribaucour1998} and \citet{Bugler2015} found the NTC region for
n-pentane to be between \SIlist{760;910}{\K} at pressures near \SI{10}{\atm}. As we will compare
with our MV data at \SI{30}{\bar}, we note that increasing the pressure tends to shift the NTC to
higher temperatures, as mentioned previously \cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
behavior for n-pentanol in the range of \SIrange{770}{900}{\K} at \SI{30}{\bar}. In this study, we
find the NTC window for MV at \SI{30}{\bar} to be between \SIlist{720;800}{\K}. Therefore, it
appears that the methyl ester functional group causes the NTC range to occur at lower temperature as
compared to alkanes and alcohols with similar alkyl chain lengths. This result was also noted by
\citet{Hadj-Ali2009} for methyl hexanoate as the fuel.

\subsection{Pressure Traces}\label{sec:pressure-traces}

\Cref{fig:pressure-traces}a shows the pressure traces for selected experiments at \(\phi=1.0\),
\(P_C =\SI[number-unit-product={\ }]{30}{\bar}\). The three reactive pressure traces shown are at
the low-temperature end of the NTC (blue, \SI{700}{\K}), one case with two-stage ignition (orange,
\SI{733}{\K}), and one case near the high-temperature limit of the NTC region (green, \SI{774}{\K}).
Also shown is the non-reactive pressure trace for the \SI{700}{\K} case (red). By comparing the
\SI{700}{\K} pressure trace with the non-reactive pressure trace, it can be seen that there is
substantial heat release prior to main ignition as measured by the deviation of the reactive
pressure trace from the non-reactive trace. However, there is only one peak in the time derivative
of the pressure, so no first-stage ignition delay is defined for this case. It can also be seen in
\cref{fig:pressure-traces}a that the \SI{774}{\K} case shows some heat release prior to ignition,
although again there is only one peak in the time derivative of the pressure. Furthermore, the heat
release at \SI{774}{\K} appears to be more gradual than at \SI{700}{\K}.

\begin{center}
    \captionsetup{type=figure}
    \includegraphics[width=\textwidth]{figures/pressure-traces.pdf}
    \caption{Selected pressure traces around the NTC region of ignition delay.
    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\) b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\)}
    \label{fig:pressure-traces}
\end{center}

A similar trend can be observed in \cref{fig:pressure-traces}b for \(\phi=2.0\) at
\(P_C=\SI{15}{\bar}\), where pressure traces at several points around the NTC region are plotted. As
in \cref{fig:pressure-traces}a, the three reactive pressure traces shown are at the low-temperature
end of the NTC (blue, \SI{698}{\K}), one case with two-stage ignition (orange, \SI{720}{\K}), and
one case near the high-temperature limit of the NTC region (green, \SI{782}{\K}). Also shown is the
non-reactive pressure trace for the \SI{698}{\K} case (red). As for the \(\phi=1.0\) case, the
pressure traces show significant heat release prior to the overall ignition, as judged by deviation
from the non-reactive case.

\section{Computational Results}\label{sec:computational-results}

\Cref{fig:simulation-comparison} compares experimentally measured overall ignition delays with
ignition delays computed with the detailed model of \citet{Dievart2013} (solid lines).
\Cref{fig:simulation-comparison}a shows results at \(P_C = \SI{15}{\bar}\), while
\cref{fig:simulation-comparison}b shows results at \(P_C = \SI{30}{\bar}\). Only some equivalence
ratios are shown for each pressure condition; data and simulated results are not shown for cases
where the reactive simulated temperature at the EOC deviated substantially from the non-reactive
temperature due to heat release during the compression stroke. Furthermore, it is important to note
that the model of \citet{Dievart2013} was not validated for MV ignition delays, only for extinction
strain rates.

\begin{center}
    \captionsetup{type=figure}
    \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
    \caption{Comparison of experimental and simulated results. a) \SI{15}{\bar}; b) \SI{30}{\bar}}
    \label{fig:simulation-comparison}
\end{center}

At \SI{15}{\bar}, the experimental ignition delays are under-predicted by the \citet{Dievart2013}
model for the three equivalence ratios shown. For the \(\phi =\) \numlist{0.25;0.5} conditions, the
model appears to be predicting an NTC region of the ignition delays as the temperature decreases,
although such a trend is not observed in the data. However, at \(\phi = 2.0\), the model does not
predict the presence of an NTC region, although one is present in the experiments. Nonetheless, the
agreement seems to be improving as the temperature is decreased. Comparing the \citet{Dievart2013}
model to the stoichiometric data at \SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\
P_C=\SI{15}{\bar}\) data. The model does not predict the NTC region found experimentally for the
\(\phi = 1.0\) experiments, but the agreement improves as the temperature decreases. Interestingly,
two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
delay. While the model of \citet{Dievart2013} over-predicts the first-stage ignition delay, it also
over-predicts the first-stage pressure rise, thereby leading to rapid overall ignition.

To elucidate the underlying reasons for the disagreement between the \citet{Dievart2013} model and
the data, we constructed an additional model using RMG (see
\cref{sec:reaction-mechanism-generator}). As can be seen in \cref{fig:simulation-comparison}a, the
agreement between the RMG model (dashed lines) and the experimental data is similar to the
\citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays, and does not
predict the NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is
similar to the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature
ignition delays and does not predict the experimental NTC region. Finally, as in the
\citet{Dievart2013} model, two-stage ignition is predicted for all of the \(\phi=1.0\) and
\(\phi=2.0\) data shown in \cref{fig:simulation-comparison}. However, the first-stage ignition
delays are \SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
delay. Again, the over-prediction of the first-stage ignition delay and pressure rise requires
further investigation and understanding of the complex low-temperature chemistry.

It is clear that neither model is able to predict the ignition delays of MV particularly well. In
addition to the poor agreement shown in \cref{fig:simulation-comparison}, the simulations for
\(P_C = \SI{15}{\bar}\), \(\phi=1.0\) and \(P_C = \SI{30}{\bar}\), \(\phi=\) \numlist{0.25;0.5;2.0}
showed substantial heat release during the compression stroke (i.e., the simulations are much too
reactive), and so these conditions aren't compared in \cref{fig:simulation-comparison}. We note
again that the model by \citet{Dievart2013} was validated for MV combustion only by comparison to
flame extinction limits, so the disagreement is not wholly surprising.

In general, there could be three likely sources of error in the models: missing reaction pathways,
incorrect values of the reaction rates, and incorrect values for thermodynamic properties of the
species. We have noted in \cref{sec:reaction-mechanism-generator} that the RMG model has many more
reactions than the \citet{Dievart2013} model and the algorithm used in RMG considers a substantial
number of the possible pathways. This reduces the possibility of missing reaction pathways affecting
the model. Further detailed studies are required to ensure that the RMG model includes all of the
relevant reaction pathways, which are outside the scope of this work.

The second source of error may be incorrect reaction rate parameters, either because the rates are
specified incorrectly in the model or because the rates are not well estimated by the
typical analogy based-rules. It should be noted that errors of this type may affect the model
generated by RMG---if the rates are not estimated correctly, reactions that are important in reality
may not be included in the model. Determining the accuracy of the reaction rates used in the RMG and
\citet{Dievart2013} models requires further detailed studies of the models, which are also outside
the scope of this work. Another, related, source of error could be incorrect estimation of the
pressure dependence of the reaction rates, which may be particularly important for the isomerization
reactions prevalent in low-temperature chemistry.

The third source of error may lie in the estimation of the thermodynamic properties of the species,
particularly the fuel radicals. In the work of \citet{Dievart2013}, the program THERM
\cite{Ritter1991} is used to estimate thermodynamic values using the group additivity method. In the
RMG model constructed in this work, RMG itself estimates the thermodynamic properties of the
molecules also using the group additivity method. Nonetheless, the two models have differing
predictions of the thermodynamic properties of the species in the model, particularly the fuel and
its radicals. The values of the heats of formation of the fuel and its \ce{H}-atom abstraction
radicals are shown in \cref{tab:heat-of-formation}; the radicals are labeled according to the
convention shown in \cref{fig:mv-structure}.

\begin{center}
    \captionsetup{type=figure}
    \includegraphics[width=60mm]{figures/Methyl_pentanoate.png}
    \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:heat-of-formation} and \cref{tab:mv-radicals}}
    \label{fig:mv-structure}
\end{center}

\begin{center}
    \captionsetup{type=table}
    \caption{Heats of formation of MV and its radicals, labeled according to the convention used in \cref{fig:mv-structure}}
    \label{tab:heat-of-formation}
    \begin{tabular}{cSSSS}
        \toprule
        \multirow{2}{*}{Radical Site} & \multicolumn{2}{c}{\citet{Dievart2013}} & \multicolumn{2}{c}{RMG Model (this work)} \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5}
        & {[\si{\kJ\per\mole}]} & {[\si{\kilo\calorie\per\mole}]} & {[\si{\kJ\per\mole}]} & {[\si{\kilo\calorie\per\mole}]} \\
        \midrule
        MV & -470.98 & -112.57 & -472.53 & -112.94 \\
        2 & -297.16 & -71.02 & -273.63 & -65.40 \\
        3 & -277.03 & -66.21 & -273.63 & -65.40 \\
        4 & -277.03 & -66.21 & -278.61 & -66.59 \\
        5 & -265.94 & -63.56 & -267.53 & -63.94 \\
        M & -270.51 & -64.65 & -270.12 & -64.56 \\
        \bottomrule
    \end{tabular}
\end{center}

\Cref{tab:heat-of-formation} shows that the heats of formation of the fuel and radicals 3, 4, 5, and
M are quite similar between the two mechanisms. However, the heat of formation of the second
radical, the one closest to the methyl ester group, has a significantly lower heat of formation in
the model by \citet{Dievart2013} than in the RMG model. Note that it is expected that the second
radical will be somewhat more stable than the other radicals, due to the influence of the methyl
ester group on the adjacent carbon atom.

This difference in heats of formation affects the pathways that consume the fuel. By conducting a
reaction pathway analysis to determine which radicals are formed from the breakdown of the fuel, we
can analyze the proportion of each radical formed as the fuel breaks down during the autoignition
process. The following analysis is conducted for a constant volume, adiabatic simulation with
initial temperature and pressure of \SI{700}{\K}, \SI{30}{\bar}, and for the stoichiometric
equivalence ratio. The rates of production of the species have been integrated until the time of
\SI{20}{\percent} fuel consumption. The results of this analysis are shown in \cref{tab:mv-radicals}
for the two models. The percentages shown in the \cref{tab:mv-radicals} are the percent of the fuel
consumed to form a particular fuel radical by all the reactions that can form that radical, and the
radicals are labeled according to the convention in \cref{fig:mv-structure}.

\begin{center}
    \captionsetup{type=table}
    \caption{Percent of MV consumed to form fuel radical species with a hydrogen atom missing at
    the location indicated in the first column and \cref{fig:mv-structure}}
    \label{tab:mv-radicals}
    \begin{tabular}{cSSS}
        \toprule
        Radical Site & {\citet{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} & {RMG switched [\si{\percent}]}\\
        \midrule
        2 & 29.2 & 12.5 & 11.0 \\
        3 & 17.5 & 12.2 & 11.1 \\
        4 & 17.5 & 50.6 & 56.6 \\
        5 & 9.5 & 3.9 & 4.3 \\
        M & 26.3 & 20.8 & 16.9 \\
        \bottomrule
    \end{tabular}
\end{center}

At the relatively low temperature and high pressure condition of this analysis, all of the fuel is
consumed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
models have quite different distributions of products from the first \ce{H}-abstraction reactions.
The model of \cite{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
prevalent, while the RMG model predicts that the radical on the fourth carbon in the chain will be
primarily formed. This is in line with the heats of formation in \cref{tab:heat-of-formation}, where
the most stable radical (i.e., the radical with the smallest heat of formation) is most likely to be
formed in each model.

To further compare the models with each other, the NASA polynomials representing the thermodynamic
properties of MV and the 5 fuel radicals from the model of \citet{Dievart2013} and are used to
replace the equivalent molecules in the RMG model. The results of a path analysis at the same
condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switched'' column.
This analysis shows that the radical on the fourth carbon atom is still the most prevalant, despite
changing the heats of formation of the fuel and its radicals.

Taken together, these results show that the poor performance in a given model cannot be attributed
to a single source. There is a strong interaction between the thermodynamics of the species and the
kinetics of the reactions, requiring further detailed study of the methyl ester system to accurately
predict the low temperature ignition delays of methyl valerate.

\section{Conclusions}\label{sec:conclusions}

In this study, we have measured ignition delays for methyl valerate over a wide range of
engine-relevant pressures, temperatures, and equivalence ratios. An NTC region of the ignition delay
and two-stage ignition are recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\) and \SI{30}{\bar}
at \(\phi=1.0\). A detailed chemical kinetic model available in the literature is unable to
reproduce the experimental results, so a new model is constructed using the Reaction Mechanism
Generator software. Although the new model contains many more reactions than the literature model,
it is still unable to predict the experimental ignition delays satisfactorily. Both models predict
an NTC region of the ignition delay under conditions where none is found in the experiments, and
fail to predict the NTC region of ignition delay that is present in the experiments. Possible
reasons for the discrepancy include missing reaction pathways, incorrect rate estimates, and
incorrect thermodynamic property estimates. Comparative analysis of the two models failed to
identify a single source of the error, and further detailed studies are required to improve
predictions of the ignition delay at these engine-relevant conditions.

\section*{References}
\bibliography{methyl-valerate}

\end{document}